---
title: "R Notebook"
output: html_notebook
---

This is a demo notebook for accessing/processing NASA data from the GES DISC.

The basic workflow is 

* Make CMR requests to search for specific products
* Obtain Amazon S3 credentials
* Access desired datasets 
* Perform analysis and display results

*** 

### Search CMR Catalogs and Obtain S3 URLs

This step uses the `httr2` package and the CMR catalog to search for data products

```{r cmr_libs, include=TRUE}
library(httr2)
library(tidyr)
library(purrr)
library(stringr)
library(s3fs)
library(ncdf4)
```

Make the CMR API request for OCO-2 LtCO2 products for Version 11.1r

The final object parses the JSON response with the product information

```{r cmr_req, include=TRUE}

cmrurl <- 'https://cmr.earthdata.nasa.gov/search/granules'

short_name <- 'OCO2_L2_Lite_FP'
start_time <- '2020-07-05T00:00:00Z'
end_time <- '2020-07-07T00:00:00Z'
tmstr <- paste(start_time,end_time,sep=",")

rqoco <- request(cmrurl) %>% 
         req_headers("Accept" = "application/json") %>% 
         req_url_query(short_name=short_name,temporal=tmstr,page_size="200",version='11.1r') %>%
         req_method("GET") %>% req_perform()     

granv11 <- rqoco %>% resp_body_json()

```

Now, the S3 URLs of the relevant products are extracted 

```{r cmr_s3_list, include=TRUE}

nprd <- length(granv11$feed$entry)
s3url <- NULL

hrefs3 <- function(lnkidx, rflst) {
    # Extract the full link from a GES DISC JSON link list
    lnkvl <- rflst[[lnkidx]]$href
    return(lnkvl)
}

for (i in seq(1,nprd)) {
    nlnk <- length(granv11$feed$entry[[i]]$links)
    mp1 <- unlist(map(seq(1,nlnk), .f= hrefs3, rflst = granv11$feed$entry[[1]]$links))
    mtchs3 <- str_detect(mp1,"s3://")
    s3url <- c(s3url,mp1[mtchs3])
}

print(s3url)
```


*** 

### Obtain Amazon S3 credentials

Instead of using a .netrc file, here the Earthdata credentials are input with the `req_auth_basic` function in the `httr2` package. A prompt for the password appears for the first run. The credentials can be archived in the `.Rhistory`

```{r s3_cred, include=TRUE}

gesdisc_s3 <- "https://data.gesdisc.earthdata.nasa.gov/s3credentials"

# Get S3 credentials
rqs3 <- request(gesdisc_s3) %>% req_auth_basic(username = 'jonhobbs515') %>% req_perform() 
if (rqs3$status_code == 200) {
  s3crd <- rqs3 %>% resp_body_json(check_type = FALSE)
}


```

### Access desired datasets

S3 direct access 
```{r s3_access, include=TRUE}
# Set access environmental paramters

s3acc <- S3FileSystem$new(aws_access_key_id = s3crd$accessKeyId,
                          aws_secret_access_key = s3crd$secretAccessKey,
                          aws_session_token = s3crd$sessionToken,
                          region_name = 'us-west-2')
s3acc$file_info(s3url[1])

# ncdf4 built without zarr support?
nc1 <- nc_open(s3url[1])
ltsdg <- ncvar_get(nc1,"sounding_id")
nc_close(nc1)


```


