---
title: "OCO-2 Data Access and Quick Aggregation"
output:
  github_document:
    df_print: paged
---

### Author: Jon Hobbs
### Date Authored: 2024-08-31

**Testing the earthdatalogin package for S3 access on AWS**

```{r libraries, echo=TRUE, results='hide'}
suppressPackageStartupMessages({ 
    library(httr)
    library(tidyr)
    library(dplyr)
    library(ggplot2)
    library(viridisLite)
    library(earthdatalogin)
    library(purrr)
    library(stringr)
    library(aws.s3)
    library(hdf5r)
})


```


Obtain S3 credentials


```{r s3_token, include=TRUE}
edl_s3_token(daac = "https://data.gesdisc.earthdata.nasa.gov")
```

## Search CMR Catalogs and Obtain S3 URLs

* Make the CMR API request for OCO-2 LtCO2 products for Version 11.1r
* The final object parses the JSON response with the product information
* `httr` R package has similar capability to the Python `requests` library
* Final object has JSON response parsed

```{r cmr_req, include=TRUE}

cmrurl <- 'https://cmr.earthdata.nasa.gov/search/granules'

short_name <- 'OCO2_L2_Lite_FP'
start_time <- '2020-07-05T00:00:00Z'
end_time <- '2020-07-07T00:00:00Z'
tmstr <- paste(start_time,end_time,sep=",")

cmrqry <- list(short_name=short_name,temporal=tmstr,page_size="200",version="11.1r")
resp <- GET(cmrurl,query=cmrqry)
http_type(resp)

granv11 <- content(resp,as="parsed") 

```

Now, the S3 URLs of the relevant products are extracted 

```{r cmr_s3_list, include=TRUE}

nprd <- length(granv11$feed$entry)
s3url <- NULL

hrefs3 <- function(lnkidx, rflst) {
    # Extract the full link from a GES DISC JSON link list
    lnkvl <- rflst[[lnkidx]]$href
    return(lnkvl)
}

for (i in seq(1,nprd)) {
    nlnk <- length(granv11$feed$entry[[i]]$links)
    mp1 <- unlist(map(seq(1,nlnk), .f= hrefs3, rflst = granv11$feed$entry[[1]]$links))
    mtchs3 <- str_detect(mp1,"s3://")
    s3url <- c(s3url,mp1[mtchs3])
}

print(s3url)
```

## Access desired datasets

* S3 direct access enabled by `aws.s3` package
* Check options for the s3fs opjects, e.g. open S3 file object before accessing NetCDF

```{r s3_access, include=TRUE}
# S3 access dataset
Sys.setenv("AWS_DEFAULT_REGION" = "us-west-2")
hf <- s3read_using(FUN = H5File$new, object = s3url[1])
hf$ls(recursive=FALSE)$name
lat <- hf[['latitude']][]
lon <- hf[['longitude']][]
qflg <- hf[['xco2_quality_flag']][]
ltxco2 <- hf[['xco2']][]
ltsdg <- hf[['sounding_id']][]
hf$close_all()

table(qflg)
lttbl <- tibble(SoundingID=as.vector(ltsdg), Latitude=as.vector(lat), Longitude=as.vector(lon),
                XCO2=as.vector(ltxco2), V11QFlag=as.vector(qflg))
lttbl <- lttbl %>% mutate(Sdg10s = floor(SoundingID / 1.0e3))
lttbl <- lttbl %>% filter(V11QFlag == 0)

```

```{r ltco2_frm, include=TRUE}
print(head(lttbl))
```

### Group and Summarize

```{r grpsmry, include=TRUE}

# Use the pivot_longer, pivot_wider approach
ltlng <- lttbl %>% pivot_longer(cols = c("XCO2","Latitude","Longitude"), names_to = "GeoVar",
                             names_prefix="", values_to = "value")
ltgrp <- ltlng %>% group_by(Sdg10s,GeoVar) %>% summarise(Med = median(value,na.rm=TRUE), NSmp=n()) %>% ungroup()
ltgrp <- ltgrp %>% filter(NSmp > 30)
ltwdr <- ltgrp %>% pivot_wider(id_cols = c("Sdg10s"), names_from=c("GeoVar"), names_prefix=c("Med_"), 
                               values_from="Med")

```


```{r map10s, include=TRUE}

# Map results
suppressPackageStartupMessages({ 
    library("rnaturalearth")
    library("rnaturalearthdata")
    library(sf)
})


theme_mat = theme_bw() 
theme_mat$axis.title.x$size = 11
theme_mat$axis.text.x$size = 10
theme_mat$axis.title.y$size = 11
theme_mat$axis.text.y$size = 10
theme_mat$plot.title$size = 12
theme_mat$legend.position = "right"
theme_mat$plot.title$hjust = 0.5
theme_mat$strip.text$size = 11
theme_mat$legend.text$size = 10
theme_mat$legend.title$size = 11

mrg_fnl <- as.data.frame(ltwdr)
mrg_fnl_sf <- st_as_sf(mrg_fnl, coords = c("Med_Longitude", "Med_Latitude"), 
                        crs = 4326, agr = "constant")

world <- ne_coastline(scale = "medium", returnclass = "sf")
p9 = viridis(9)

ltmap <- ggplot(data = world) + geom_sf(color = '#777777') + 
  geom_sf(data = mrg_fnl_sf,aes(color=Med_XCO2), size=0.8) +
  scale_color_gradientn("XCO2 [ppm]",colors=p9)  + 
  theme_mat +  ggtitle("OCO-2 10s Median XCO2")  

ltmap
```